The Lasting Shadows of Economic Downturns

### 2024-02-14 outputs/paper/paper.qmd
Update paper.qmd

Second, we obtained unemployment data for the years 1976-1989 which Stuart obtained through email communication with Brian Hannon of the BLS from the Bureau of Labor Statistics (BLS) Local Area Unemployment Statistics [@BLS]. This data, covering various counties, is compiled in the file 'BLS_all_archived_counties7689aa (UNOFFICIAL).xlsx' and is available through the Inter-university Consortium for Political and Social Research.

@@

# Discussion
## How Recession Impacts on Unemployment
## How Recession Impacts on Personal Income
## How Recession Changes the Business Patterns
## Limitations and Next Steps
Due to limitations in accessing relevant datasets, this paper will not explore the diverse social impacts of recessions, specifically on birth rates and education, which were initially of primary interest. Drawing from Stuart's foundational research, we quantitatively assess the prolonged impacts of the 1980–1982 recession on individuals who were in their formative years—children, adolescents, and young adults—at the onset of this economic downturn. Stuart's findings reveal significant long-term detriments to educational attainment and income levels. Specifically, for individuals aged 0–10 in 1979, a 10 percent reduction in per capita earnings from 1979 to 1982 within their birth county correlates with a 4.7 percentage point (14.6 percent) reduction in the attainment of four-year college degrees, a decrease in earned income by $2,100 (5.2 percent), and a 1.6 percentage point (13.0 percent) increase in the likelihood of living in poverty during 2000–2013. The adverse effects on college graduation rates are notably pronounced and consistently significant for individuals aged 0–13 in 1979, and diminish to small and statistically non-significant levels for those aged 14–22. This pattern indicates that the recession's impact likely stems from a reduction in childhood human capital or a sustained decrease in parental financial capacity to support higher education. Future research will address the effects on education and birth rates once relevant and accessible datasets become available.

>>>>

Second, we obtained unemployment data for the years 1976-1989 which Stuart obtained from the Bureau of Labor Statistics (BLS) Local Area Unemployment Statistics [@BLS] through email communication with Brian Hannon. This data, covering various counties, is compiled in the file 'BLS_all_archived_counties7689aa (UNOFFICIAL).xlsx' and is available through the Inter-university Consortium for Political and Social Research.

@@

# Discussion
## How Recession Impacts on Unemployment
## How Recession Impacts on Personal Income
## How Recession Changes the Business Patterns

## Limitations and Next Steps
Due to the unavailability of datasets, this paper will not explore the further social consequences of economic recessions specifically focusing on how they affect birth rates and education which were initially the main areas of interest. Drawing inspiration from Stuart's study [@10.1257/app.20180055], we quantitatively evaluate the long-term effects of the recession that occurred between 1980 and 1982 on individuals who were in their years—children, adolescents, and young adults—when this economic downturn began. Stuart's findings reveal lasting impacts on educational achievements and income levels. For instance among individuals aged 0–10 in 1979, a decrease of 10 percent in per capita earnings from 1979 to 1982 within their birth county is associated with a decline of 4.7 percentage points in four-year college degree attainment. Additionally, it leads to a drop in earned income by $2,100 as an increase of 1.6 percentage points in the likelihood of experiencing poverty during the period from 2000 to 2013. The negative impact on college graduation rates is particularly notable and consistently significant for individuals aged 0–13 in 1979. Becomes smaller and statistically insignificant for those between 14 and 22 years old. This trend indicates that the recession's influence likely originates from a reduction in childhood capital or sustained financial constraints, among parents affecting their ability to support education. In the future when we have access, to easily accessible datasets researchers will investigate how education and birth rates are impacted.
